### 2021-08-17 paper/paper-misspec-supp.tex
finish later

%% LyX 1.3 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[english, 12pt]{article}
\usepackage{times}
%\usepackage{algorithm2e}
\usepackage{url}
\usepackage{bbm}
\usepackage[T1]{fontenc}
\usepackage[latin1]{inputenc}
\usepackage{geometry}
\geometry{verbose,letterpaper,tmargin=2cm,bmargin=2cm,lmargin=1.5cm,rmargin=1.5cm}
\usepackage{rotating}
\usepackage{color}
\usepackage{graphicx}
\usepackage{amsmath, amsthm, amssymb}
\usepackage{setspace}
\usepackage{lineno}
\usepackage{hyperref}
\usepackage{bbm}
\usepackage{makecell}
\usepackage{placeins}
\usepackage{subcaption}

%\renewcommand{\arraystretch}{1.8}

%\linenumbers
%\doublespacing
\onehalfspacing
%\usepackage[authoryear]{natbib}
\usepackage{natbib} \bibpunct{(}{)}{;}{author-year}{}{,}

%Pour les rajouts
\usepackage{color}
\definecolor{trustcolor}{rgb}{0,0,1}

\usepackage{dsfont}
\usepackage[warn]{textcomp}
\usepackage{adjustbox}
\usepackage{multirow}
\usepackage{subcaption}
\usepackage{graphicx}
\graphicspath{{../figures/}}
\DeclareMathOperator*{\argmin}{\arg\!\min}

\let\tabbeg\tabular
\let\tabend\endtabular
\renewenvironment{tabular}{\begin{adjustbox}{max width=0.95\textwidth}\tabbeg}{\tabend\end{adjustbox}}

\makeatletter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
%% Bold symbol macro for standard LaTeX users
%\newcommand{\boldsymbol}[1]{\mbox{\boldmath $#1$}}

%% Because html converters don't know tabularnewline
\providecommand{\tabularnewline}{\\}
\renewcommand*{\arraystretch}{1.2}

\usepackage{babel}
\makeatother

\begin{document}

\renewcommand{\thefigure}{S\arabic{figure}}
\setcounter{figure}{0}
\renewcommand{\thetable}{S\arabic{table}}
\setcounter{table}{0}
\renewcommand{\theequation}{S\arabic{equation}}
\setcounter{equation}{0}

\section*{Supplementary Materials}

\vspace*{5em}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\FloatBarrier

\begin{figure}[h]
	\centerline{\includegraphics[width=0.8\textwidth]{simu-qc-plot}}
	\caption{Quality control plot, as proposed in \cite{prive2020ldpred2}, for the simulations with sample size misspecification. The standard deviations are derived from the summary statistics assuming the same global GWAS sample size for all variants (300,000).}
	\label{fig:qcplot}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\FloatBarrier

\begin{figure}[p]
	\begin{subfigure}{\textwidth}
		\centering
		\includegraphics[width=.8\linewidth]{lassosum2-mis-maxN-corr1}
	\end{subfigure}\vspace{2em}
	\begin{subfigure}{\textwidth}
		\centering
		\includegraphics[width=.8\linewidth]{lassosum2-mis-imputeN-corr1}
	\end{subfigure}
	\caption{Multiple results (squared correlation r2 between polygenic score and phenotype) over a grid of parameters ($\lambda$ and $\delta$) for lassosum2 in one of the simulations with misspecified GWAS sample sizes (N). The top subfigure is using the maximum N, whereas the bottom subfigure is using the imputed N.  Missing points represent models that were detected as divergent.}
	\label{fig:lassosum2-misN}
\end{figure}

\begin{figure}[p]
	\begin{subfigure}{\textwidth}
		\centering
		\includegraphics[width=.8\linewidth]{ldpred2-mis-maxN-corr1}
	\end{subfigure}\vspace{2em}
	\begin{subfigure}{\textwidth}
		\centering
		\includegraphics[width=.8\linewidth]{ldpred2-mis-imputeN-corr1}
	\end{subfigure}
	\caption{Multiple results (squared correlation r2 between polygenic score and phenotype) over a grid of parameters (the SNP heritability h2 and the proportion of causal variants p) for LDpred2(-grid) in one of the simulations with misspecified GWAS sample sizes (N). The top subfigure is using the maximum N, whereas the bottom subfigure is using the imputed N. Note that the small h2 values (0.002, 0.02, 0.06) were added and are part of the method we call ``LDpred2-low-h2'' in the main text. Missing points represent models that were detected as divergent.}
	\label{fig:ldpred2-misN}
\end{figure}

\begin{figure}[p]
	\begin{subfigure}{\textwidth}
		\centering
		\includegraphics[width=.8\linewidth]{ldpred2-mis-maxN-corr2}
	\end{subfigure}\vspace{2em}
	\begin{subfigure}{\textwidth}
		\centering
		\includegraphics[width=.8\linewidth]{ldpred2-mis-imputeN-corr2}
	\end{subfigure}
	\caption{Same results as figure \ref{fig:ldpred2-misN}, but using the LD with independent blocks.}
	\label{fig:ldpred2-misN-corr2}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\FloatBarrier

\begin{figure}[p]
	\centerline{\includegraphics[width=0.8\textwidth]{compare-sd}}
	\caption{Standard deviations from ``true'' genotype calls versus from imputed dosages, colored by INFO scores. \label{fig:compare-sd}}
\end{figure}

\begin{figure}[p]
	\centerline{\includegraphics[width=0.8\textwidth]{compare-beta}}
	\caption{GWAS effect sizes from ``true'' genotype calls versus from imputed dosages, colored by INFO scores. \label{fig:compare-beta}}
\end{figure}

\begin{figure}[p]
	\centerline{\includegraphics[width=0.8\textwidth]{compare-beta-se}}
	\caption{Standard errors of GWAS effect sizes from ``true'' genotype calls versus from imputed dosages, colored by INFO scores. \label{fig:compare-beta-se}}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure}[p]
	\centerline{\includegraphics[width=0.8\textwidth]{compare-info}}
	\caption{INFO scores reported in the UK Biobank vs.\ INFO scores recomputed from the subset of 362,307 European individuals used in this paper. \label{fig:compare-info}}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\FloatBarrier

\begin{figure}[p]
	\centerline{\includegraphics[width=0.8\textwidth]{compare-mult-imp}}
	\caption{GWAS effect sizes and Z-scores from imputed dosages versus from multiple imputation (MI, 10 draws used), ``corrected'' and colored by INFO scores. This shows only one tenth of the simulated variants, because MI takes some time to run. \label{fig:compare-mi}}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\FloatBarrier

\begin{figure}[p]
	\centerline{\includegraphics[width=0.65\textwidth]{brca_icogs_hist_info}}
	\caption{Histogram of the imputation INFO scores of the HapMap3 variants from the iCOGS GWAS summary statistics for breast cancer \cite[]{michailidou2013large}. \label{fig:hist-info-brca2}}
\end{figure}

\begin{figure}[p]
	\centerline{\includegraphics[width=0.65\textwidth]{brca_onco_hist_info}}
	\caption{Histogram of the imputation INFO scores of the HapMap3 variants from the OncoArray GWAS summary statistics for breast cancer \cite[]{michailidou2017association}. [TODO: REEXPORT -- WAS CORRUPTED] \label{fig:hist-info-brca}}
\end{figure}

\FloatBarrier

\begin{figure}[p]
	\centerline{\includegraphics[width=0.8\textwidth]{hist_bad_pos}}
	\caption{Histogram of the positions of outlier variants from Figure 2B [TODO: VERIF NUMBER AT THE END] in the main text. \label{fig:hist-bad-pos}}
\end{figure}

\begin{figure}[p]
	\centerline{\includegraphics[width=0.95\textwidth]{gwas_bad_pc}}
	\caption{Standard deviations inferred from the simulated GWAS summary statistics (\textbf{A:} with no covariate; \textbf{B:} with PC19 from the UK Biobank as covariate) versus the ones inferred from the allele frequencies. Only HapMap3 variants from chromosome 6 are represented.}
	\label{fig:gwas_bad_pc}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\FloatBarrier

\begin{figure}[p]
	\centerline{\includegraphics[width=0.95\textwidth]{brca_icogs_qc}}
	\caption{Standard deviations inferred from the iCOGS breast cancer GWAS summary statistics versus the ones inferred from the reported GWAS allele frequencies. Only 100,000 variants are represented, at random.}
	\label{fig:brca_icogs_qc}
\end{figure}

\begin{figure}[p]
	\centerline{\includegraphics[width=0.95\textwidth]{cad_qc}}
	\caption{Standard deviations inferred from the CAD GWAS summary statistics versus the ones inferred from the reported GWAS allele frequencies. Only 100,000 variants are represented, at random.}
	\label{fig:cad_qc}
\end{figure}

\begin{figure}[p]
	\centerline{\includegraphics[width=0.95\textwidth]{mdd_qc}}
	\caption{Standard deviations inferred from the MDD GWAS summary statistics versus the ones inferred from the reported GWAS allele frequencies. Only 100,000 variants are represented, at random.}
	\label{fig:mdd_qc}
\end{figure}

\begin{figure}[p]
	\centerline{\includegraphics[width=0.95\textwidth]{prca_qc}}
	\caption{Standard deviations inferred from the PRCA GWAS summary statistics versus the ones inferred from the reported GWAS allele frequencies. Only 100,000 variants are represented, at random. [TODO: REEXPORT FIGURE]}
	\label{fig:prca_qc}
\end{figure}

\FloatBarrier

\begin{figure}[p]
	\centerline{\includegraphics[width=0.95\textwidth]{t1d_affy_qc}}
	\caption{Standard deviations inferred from the T1D (Affymetrix) GWAS summary statistics versus the ones inferred from the reported GWAS allele frequencies. Only 100,000 variants are represented, at random.}
	\label{fig:t1d_affy_qc}
\end{figure}

\begin{figure}[p]
\centerline{\includegraphics[width=0.95\textwidth]{t1d_illu_qc}}
\caption{Standard deviations inferred from the T1D (Illumina) GWAS summary statistics versus the ones inferred from the reported GWAS allele frequencies. Only 100,000 variants are represented, at random.}
\label{fig:t1d_illu_qc}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\FloatBarrier

\begin{figure}[p]
	\centerline{\includegraphics[width=0.75\textwidth]{res-cad}}
	\caption{Raw partial correlations (for \textit{all} models) for predicting CAD in the UK Biobank from external summary statistics.}
	\label{fig:res_cad}
\end{figure}

\begin{figure}[p]
	\centerline{\includegraphics[width=0.75\textwidth]{res-mdd}}
	\caption{Raw partial correlations (for \textit{all} models) for predicting MDD in the UK Biobank from external summary statistics.}
	\label{fig:res_mdd}
\end{figure}

\begin{figure}[p]
	\centerline{\includegraphics[width=0.75\textwidth]{res-prca}}
	\caption{Raw partial correlations (for \textit{all} models) for predicting PRCA in the UK Biobank from external summary statistics.}
	\label{fig:res_prca}
\end{figure}

\begin{figure}[p]
	\centerline{\includegraphics[width=0.95\textwidth]{res-t1d}}
	\caption{Raw partial correlations (for \textit{all} models) for predicting T1D in the UK Biobank from external summary statistics.}
	\label{fig:res_t1d}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\FloatBarrier

\begin{figure}[p]
	\centerline{\includegraphics[width=0.95\textwidth]{res-all-brca}}
	\caption{Raw partial correlations (for \textit{all} models and \textit{all} sets) for predicting BRCA in the UK Biobank from external summary statistics.}
	\label{fig:res_brca_all}
\end{figure}

\begin{figure}[p]
	\centerline{\includegraphics[width=0.9\textwidth]{res-all-cad}}
	\caption{Raw partial correlations (for \textit{all} models and \textit{all} sets) for predicting CAD in the UK Biobank from external summary statistics.}
	\label{fig:res_cad_all}
\end{figure}

\begin{figure}[p]
	\centerline{\includegraphics[width=0.9\textwidth]{res-all-mdd}}
	\caption{Raw partial correlations (for \textit{all} models and \textit{all} sets) for predicting MDD in the UK Biobank from external summary statistics.}
	\label{fig:res_mdd_all}
\end{figure}

\begin{figure}[p]
	\centerline{\includegraphics[width=0.9\textwidth]{res-all-prca}}
	\caption{Raw partial correlations (for \textit{all} models and \textit{all} sets) for predicting PRCA in the UK Biobank from external summary statistics.}
	\label{fig:res_prca_all}
\end{figure}

\begin{figure}[p]
	\centerline{\includegraphics[width=0.95\textwidth]{res-all-t1d}}
	\caption{Raw partial correlations (for \textit{all} models and \textit{all} sets) for predicting T1D in the UK Biobank from external summary statistics.}
	\label{fig:res_t1d_all}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\FloatBarrier

\begin{figure}[p]
	\centerline{\includegraphics[width=0.8\textwidth]{simu_hist_info}}
	\caption{Histogram of the imputation INFO scores of the 40,000 variants from chromosome 22 of the UK Biobank data used in the simulations. \label{fig:hist-info}}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\clearpage

\FloatBarrier

\begin{figure}[p]
	\centerline{\includegraphics[width=0.85\textwidth]{pseudoval}}
	\caption{For one simulation, scores from validation vs.\ scores from pseudo-validation as described in \cite{mak2017polygenic} using either correlations (the default in lassosum) or p-values when computing local false discovery rates.}
	\label{fig:pseudoval}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\FloatBarrier

\begin{figure}[p]
	\centerline{\includegraphics[width=0.95\textwidth]{compare-cor}}
	\caption{Comparing correlations between a subset of HapMap3 variants on chromosome 22, computed in three different ways: 1/ from imputed dosages; 2/ from multiple imputation (i.e.\ generating multiple complete datasets sampled according to imputation probabilities, computing correlations, and averaging results); from imputed dosages with the multiplicative correction of $\sqrt{\text{INFO}_j \cdot \text{INFO}_k}$ (for $j \neq k$).}
	\label{fig:compare-cor}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\FloatBarrier
\clearpage

\bibliographystyle{natbib}
\bibliography{refs}

\end{document}