### 2022-02-02 keep4later/new-sets.tex
rm the two additional sets

\subsection{Two new sets of variants}

We also design two larger sets of imputed variants to compare against using only HapMap3 variants for prediction.
Following \cite{prive2021high}, we first restrict to UKBB variants with MAF > 0.01 and INFO > 0.3.
We then compile frequencies and imputation INFO scores from other datasets, iPSYCH and summary statistics for breast cancer, prostate cancer, coronary artery disease, type-1 diabetes and depression \cite[]{bybjerg2020ipsych2015,michailidou2017association,schumacher2018association,nikpay2015comprehensive,censin2017childhood,wray2018genome}.
We restrict to variants with a mean INFO > 0.3 in these other datasets, and also compute the median frequency per variant.
To exclude potential mismappings in the genotyped data \cite[]{kunert2020allele} that might have propagated to the imputed data, we compare median frequencies in the external data to the ones in the UK Biobank.
As we expect these potential errors to be localized around errors in the genotype data, we apply a moving-average smoothing on the frequency differences to increase power to detect these errors and also reduce false positives.
We define the threshold (of 0.03) on these smoothed differences based on visual inspection of their histogram.
This results in an initial set of 9,394,361 variants.

We then define the two sets from this large set of variants.
One is based on clumping, using a threshold $r^2 = 0.9$ over a radius of 100 Kbp and prioritizing HapMap3 variants and larger INFO scores. This results in a set ``clump'' of 2,465,478 variants, among which there are 554,655 of the initial HapMap3 variants.
For the second set, we aim at maximizing the tagging of all the initial 9,394,361 variants, i.e. $\sum_{i \in \text{all}} \max_{j \in \text{set}}{|R_{i,j}|}$, where $R_{i,j}$ is the correlation between variants $i$ and $j$ (inspired from the alternative sensitivity of \cite{agier2016systematic}).
We design a greedy algorithm that first selects all HapMap3 variants, then adds one variant at a time, the one that maximizes the addition to this sum, until no variant can add more than 0.2.
This results in a set ``maxtag'' of 2,029,086 variants.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\FloatBarrier

\begin{figure}[p]
	\centerline{\includegraphics[width=0.95\textwidth]{res-all-brca}}
	\caption{Raw partial correlations (for \textit{all} models and \textit{all} sets) for predicting BRCA in the UK Biobank from external summary statistics. These are computed using function \texttt{pcor} of R package bigstatsr where 95\% confidence intervals are obtained through Fisher's Z-transformation.}
	\label{fig:res_brca_all}
\end{figure}

\begin{figure}[p]
	\centerline{\includegraphics[width=0.9\textwidth]{res-all-cad}}
	\caption{Raw partial correlations (for \textit{all} models and \textit{all} sets) for predicting CAD in the UK Biobank from external summary statistics. These are computed using function \texttt{pcor} of R package bigstatsr where 95\% confidence intervals are obtained through Fisher's Z-transformation.}
	\label{fig:res_cad_all}
\end{figure}

\begin{figure}[p]
	\centerline{\includegraphics[width=0.9\textwidth]{res-all-mdd}}
	\caption{Raw partial correlations (for \textit{all} models and \textit{all} sets) for predicting MDD in the UK Biobank from external summary statistics. These are computed using function \texttt{pcor} of R package bigstatsr where 95\% confidence intervals are obtained through Fisher's Z-transformation.}
	\label{fig:res_mdd_all}
\end{figure}

\begin{figure}[p]
	\centerline{\includegraphics[width=0.9\textwidth]{res-all-prca}}
	\caption{Raw partial correlations (for \textit{all} models and \textit{all} sets) for predicting PRCA in the UK Biobank from external summary statistics. These are computed using function \texttt{pcor} of R package bigstatsr where 95\% confidence intervals are obtained through Fisher's Z-transformation.}
	\label{fig:res_prca_all}
\end{figure}

\begin{figure}[p]
	\centerline{\includegraphics[width=0.95\textwidth]{res-all-t1d}}
	\caption{Raw partial correlations (for \textit{all} models and \textit{all} sets) for predicting T1D in the UK Biobank from external summary statistics. These are computed using function \texttt{pcor} of R package bigstatsr where 95\% confidence intervals are obtained through Fisher's Z-transformation.}
	\label{fig:res_t1d_all}
\end{figure}